Results of two stellar occultations by dwarf planet (1) Ceres

### 2017-08-16 Ceres/Old_versions/Altair 23.03.2015/CeresV3.tex
mod 2017 ago16

Ceres is the sole exemplar of a dwarf planet in the inner Solar System. Far from being a mere taxonomic information, this suggests the great impact its study can have on the understanding of planetary formation and evolution of the Solar System. Indeed, it was proposed that Ceres could have its origin as a transneptunian object, latter scattered to the Main Belt due to the giant planets' migration predicted by the ``Nice Model''. It may well had been formed close to its current location, however, even on this scenario, the dynamical history of the Solar System must have let its signatures on Ceres. Not only on the late heavy bombardment features that might exist on its surface, but also on its volatiles, which could have been transported from the outer region.

@@

The first stellar occultation by Ceres was observed in 1984 \citep{Millis1987} and lead to the determination of its size to the precision of the kilometre, at a time when the uncertainties were often ten times larger. The low apparent magnitude of Ceres, if compared to most asteroids, imposes a somewhat strong constraint on the stars capable of causing a detectable magnitude drop when occulted. For instance, after the 1984 event, to our knowledge, only four like this have been observed. Two of them had only two chords each, not sufficient, thus, for providing accurate measurements\footnote{These events took place on 22 August 1994 and 30 October 2010.}. The two remaining events, which occurred on 17 August 2010 and 25 October 2013, are reported on the present work. The former was observed in Brazil from 5 stations, while the later was from the United States and led to 9 chords. Throughout the paper we shall refer to these events as the ``2010'' and the ``2013 occultation''.

@@

Of course, the apparent semimajor axis $a^\prime$ equals the equatorial radius $a$ of the ellipsoid. %The polar radius $b$ can then be determined from $\epsilon = 1 - (b/a)$.

@@

Evaluation of the reduced $\chi^2$ function over the $\lbrace a, \epsilon^\prime, P, f_c, g_c \rbrace$--space produces a map related to the likelihood of a solution. For instance, the 1--sigma reported time uncertainties correspond to the subset of the parameter space for which $\chi^2_r$ lies between ${\chi^2_r}_{min}$ and ${\chi^2_r}_{min} + 1$. This allows the determination of the error bars of the physical parameters. It is worthwhile to mention that we consider all the uncertainty as due to timing issues.

@@

The resulting values of semimajor axis, oblateness, position angle and centre coordinates are presented in the second column of Table~2.%, and the corresponding ellipse is plotted with the observed chords in Fig.~2.
% The equivalent radius of the best fit solution is $R_{equiv} = 469 \pm 10$ km.

@@

We notice that the constraint corresponds to the upper limit of nominal solution's 1-sigma error bar. On the other hand, it selects the smallest values of semimajor axis, improving its determination by a factor of about 2. Notwithstanding, oblateness' figures remain the same.

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Results of limb fitting to the data of the 2010 and 2013 events.}
  \begin{tabular}{@{}lcccc}
  \hline
     Solution & 2010/Nominal & 2010/Pole-constrained & 2013/Nominal & 2013/Pole-constrained \\
\hline
Equatorial Radius (km) & 491 $\pm$ 7   & 486 $\pm$ 3   & 486 $\pm$ 2   & \\
Oblateness          & 0.08$\pm$0.03 & 0.08$\pm$0.03 & 0.09$\pm$0.05 & \\
Position angle (deg)& 5 $\pm$ 10    & 12 $\pm$ 3 (*)& 22 $\pm$ 5    &  \\
$f_c$ (km)          & 133 $\pm$ 9   & 138 $\pm$ 5   & 78 $\pm$ 7    & \\
$g_c$ (km)          & -17 $\pm$ 15  & -11 $\pm$ 11  & 14 $\pm$ 16   & \\
$\chi^2_{r,min}$    & 0.24          &  0.42         & 1.27          & \\
\hline
\end{tabular}
\textbf{Notes.} Error bars are at 1-sigma level. The reported oblateness derived from the 2010 event are the true ones, see text for details. \\
(*) Position angle derived from Ceres' rotational pole coordinates determined by \cite{Drummond2014}.
\end{minipage}
\end{table*}

@@

Nine positive chords were obtained for the 2013 occutation by a variety of equipment as listed in Table \ref{Tab: obs-2013}. Each station was equipped with a video camera (whose readout time may be considered negligible). This is of particular importance in this event, since Ceres' shadow had the much faster speed of 42.6 $km/s$, if compared to the 2010 event. All the observing sites are in the United States, as depicted in Figure \ref{Fig: Ceres-2013-map}.

@@

From the occultation times obtained, the shape for Ceres' limb was adjusted by an ellipse characterized by five parameters: the coordinates of the body center, relative to the star in the plane of the sky ($f_{c}, g_{c}$); the apparent semimajor axis a, equivalent to the equatorial radius $R_{Equa}$ ; the apparent oblateness = $(a'-b')/a'$ (where $b'$ is the apparent semiminor axis); and the position angle P of the semiminor axis $b'$ . The coordinates $f_{c}$ and $g_{c}$, in kilometers, are calculated using the JPL\#33  Ceres' ephemeris and the star position given in Section \ref{Sec: observation-2013}. They are positive toward the local celestial east and north, respectively.  The position angle P is counted positively from the direction of local celestial north to celestial east.

>>>>

Ceres is the sole exemplar of a dwarf planet in the inner Solar System. Far from being a mere taxonomic information, this suggests the great impact its study can have on the understanding of planetary formation and evolution of the Solar System. Indeed, it was proposed that Ceres could have its origin as a transneptunian object, later scattered to the Main Belt due to the giant planets' migration predicted by the ``Nice Model''. It may well had been formed close to its current location, however, even on this scenario, the dynamical history of the Solar System must have let its signatures on Ceres. Not only on the late heavy bombardment features that might exist on its surface, but also on its volatiles, which could have been transported from the outer region.

@@

The first stellar occultation by Ceres was observed in 1984 \citep{Millis1987} and lead to the determination of its size to the precision of the kilometre, at a time when the uncertainties were often ten times larger. The low apparent magnitude of Ceres, if compared to most asteroids, imposes a somewhat strong constraint on the stars capable of causing a detectable magnitude drop when occulted. For instance, after the 1984 event, to our knowledge, only four like this have been observed. Two of them had only two chords each, not sufficient, thus, for providing accurate measurements\footnote{These events took place on 22 August 1994 and 30 October 2010.}. The two remaining events, which occurred on 17 August 2010 and 25 October 2013, are reported on the present work. The former was observed in Brazil from 5 stations, while the latter was from the United States and led to 9 chords. Throughout the paper we shall refer to these events as the ``2010'' and the ``2013 occultation''.

@@

Of course, the apparent semimajor axis $a^\prime$ equals the equatorial radius $R_{equa}$ of the ellipsoid. %The polar radius $b$ can then be determined from $\epsilon = 1 - (b/a)$.

@@

Evaluation of the reduced $\chi^2$ function over the $\lbrace R_{equa}, \epsilon^\prime, P, f_c, g_c \rbrace$--space produces a map related to the likelihood of a solution. For instance, the 1--sigma reported time uncertainties correspond to the subset of the parameter space for which $\chi^2_r$ lies between ${\chi^2_r}_{min}$ and ${\chi^2_r}_{min} + 1$. This allows the determination of the error bars of the physical parameters. It is worthwhile to mention that we consider all the uncertainty as due to timing issues.

@@

The resulting values of equatorial radius, oblateness, position angle and centre coordinates are presented in the second column of Table~2.%, and the corresponding ellipse is plotted with the observed chords in Fig.~2.
% The equivalent radius of the best fit solution is $R_{equiv} = 469 \pm 10$ km.

@@

We notice that the constraint corresponds to the upper limit of nominal solution's 1-sigma error bar. On the other hand, it selects the smallest values of equatorial radius, improving its determination by a factor of about 2. Notwithstanding, oblateness' figures remain the same.

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Results of limb fitting to the data of the 2010 and 2013 events.}
  \begin{tabular}{@{}lcccc}
  \hline
     Solution & 2010/Nominal & 2010/Pole-constrained & 2013/Nominal & 2013/Pole-constrained \\
\hline
Equatorial radius (km) & 491 $\pm$ 7   & 486 $\pm$ 3   & 486 $\pm$ 2   & \\
Oblateness             & 0.08$\pm$0.03 & 0.08$\pm$0.03 & 0.09$\pm$0.05 & \\
Position angle (deg)   & 5 $\pm$ 10    & 12 $\pm$ 3 (*)& 22 $\pm$ 5    &  (*)\\
$f_c$ (km)             & 133 $\pm$ 9   & 138 $\pm$ 5   & 78 $\pm$ 7    & \\
$g_c$ (km)             & -17 $\pm$ 15  & -11 $\pm$ 11  & 14 $\pm$ 16   & \\
$\chi^2_{r,min}$       & 0.24          &  0.42         & 1.27          & \\
\hline
\end{tabular}
\textbf{Notes.} Error bars are at 1-sigma level. The reported oblateness derived from the 2010 event are the physical ones (not apparent), see text for details. \\
(*) Position angle derived from Ceres' rotational pole coordinates determined by \cite{Drummond2014}.
\end{minipage}
\end{table*}

@@

Nine positive chords were obtained for the 2013 occultation by a variety of equipment as listed in Table \ref{Tab: obs-2013}. Each station was equipped with a video camera (whose readout time may be considered negligible). This is of particular importance in this event, since Ceres' shadow had the much faster speed of 42.6 $km/s$, if compared to the 2010 event. All the observing sites are in the United States, as depicted in Figure \ref{Fig: Ceres-2013-map}.

@@

From the occultation times obtained, the shape for Ceres' limb was adjusted by an ellipse characterized by five parameters: the coordinates of the body center, relative to the star in the plane of the sky ($f_{c}, g_{c}$); the apparent semimajor axis a, equivalent to the equatorial radius $R_{Equa}$ ; the apparent oblateness = $(a'-b')/a'$ (where $b'$ is the apparent semiminor axis or Polar Radius); and the position angle P of the semiminor axis $b'$ . The coordinates $f_{c}$ and $g_{c}$, in kilometers, are calculated using the JPL\#33  Ceres' ephemeris and the star position given in Section \ref{Sec: observation-2013}. They are positive toward the local celestial east and north, respectively.  The position angle P is counted positively from the direction of local celestial north to celestial east.

### 2017-08-16 Ceres/Old_versions/Altair_25_03_2015/CeresV3.tex
mod 2017 ago16

\begin{figure}
\includegraphics[scale=0.42]{figures/Ceres_2010.png} 
\caption{Post-occultation reconstruction of Ceres' shadow path on Earth for the 2010 August 17 event. The big red dot is the geocentric closest approach at 22:41:28 UT. The small red ones represents the center of the shadow separated by one minute where the shadow moves from the left to the right. The blue dots are the sites that have observed the event. As described in text, UFSC had a negative chord.
\label{Fig: Ceres-2010-map}}
\end{figure}

@@

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Results of limb fitting to the data of the 2010 and 2013 events.}
  \begin{tabular}{@{}lcccc}
  \hline
     Solution & 2010/Nominal & 2010/Pole-constrained & 2013/Nominal & 2013/Pole-constrained \\
\hline
Equatorial radius (km) & 491 $\pm$ 7   & 486 $\pm$ 3   & 486 $\pm$ 2   & \\
Oblateness             & 0.08$\pm$0.03 & 0.08$\pm$0.03 & 0.09$\pm$0.05 & \\
Position angle (deg)   & 5 $\pm$ 10    & 12 $\pm$ 3 (*)& 22 $\pm$ 5    &  (*)\\
$f_c$ (km)             & 133 $\pm$ 9   & 138 $\pm$ 5   & 78 $\pm$ 7    & \\
$g_c$ (km)             & -17 $\pm$ 15  & -11 $\pm$ 11  & 14 $\pm$ 16   & \\
$\chi^2_{r,min}$       & 0.24          &  0.42         & 1.27          & \\
\hline
\end{tabular}
\textbf{Notes.} Error bars are at 1-sigma level. The reported oblateness derived from the 2010 event are the physical ones (not apparent), see text for details. \\
(*) Position angle derived from Ceres' rotational pole coordinates determined by \cite{Drummond2014}.
\end{minipage}
\end{table*}

@@

\begin{figure}
\includegraphics[scale=0.42]{figures/Ceres_2013.png} 
\caption{Post-occultation reconstruction of Ceres' shadow path on Earth for the 2013 October 25 event at the east coast of USA. The red dots represents the center of the shadow separated by one minute where the shadow moves from the left to the right. The geocentric closest approach at 09:43:11 UT. The blue dots are the sites that have observed the event.\label{Fig: Ceres-2013-map}}
\end{figure}

>>>>

\begin{figure}
\includegraphics[scale=0.42]{figures/Ceres_2010.png} 
\caption{Post-occultation reconstruction of Ceres' shadow path on Earth for the 2010 August 17 event. The big red dot is the geocentric closest approach at 22:40:25 UT. The small red ones represents the center of the shadow separated by one minute where the shadow moves from the left to the right. The blue dots are the sites that have observed the event. As described in text, UFSC had a negative chord.
\label{Fig: Ceres-2010-map}}
\end{figure}

@@

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Results of limb fitting to the data of the 2010 and 2013 events.}
  \begin{tabular}{@{}lcccc}
  \hline
     Solution & 2010/Nominal & 2010/Pole-constrained & 2013/Nominal & 2013/Pole-constrained \\
\hline
Equatorial radius (km) & 491 $\pm$ 7   & 486 $\pm$ 3   & 486 $\pm$ 2   & 485 $\pm$ 4\\
Oblateness             & 0.08 $\pm$ 0.03 & 0.08 $\pm$ 0.03 & 0.09 $\pm$ 0.05 & 0.08 $\pm$ 0.04\\
Position angle (deg)   & 5 $\pm$ 10    & 12 $\pm$ 3 (*)& 22 $\pm$ 5    &  25$\pm$3(*)\\
$f_c$ (km)             & 133 $\pm$ 9   & 138 $\pm$ 5   & 78 $\pm$ 7    & 78 $\pm$ 6\\
$g_c$ (km)             & -17 $\pm$ 15  & -11 $\pm$ 11  & 14 $\pm$ 16   & 13 $\pm$ 16\\
$\chi^2_{r,min}$       & 0.24          &  0.42         & 1.27          & 1.27\\
\hline
\end{tabular}
\textbf{Notes.} Error bars are at 1-sigma level. The reported oblateness derived from the 2010 event are the physical ones (not apparent), see text for details. \\
(*) Position angle derived from Ceres' rotational pole coordinates determined by \cite{Drummond2014}.
\end{minipage}
\end{table*}

@@

\begin{figure}
\includegraphics[scale=0.42]{figures/Ceres_2013.png} 
\caption{Post-occultation reconstruction of Ceres' shadow path on Earth for the 2013 October 25 event at the east coast of USA. The red dots represents the center of the shadow separated by one minute where the shadow moves from the left to the right. The geocentric closest approach at 09:43:03 UT. The blue dots are the sites that have observed the event.\label{Fig: Ceres-2013-map}}
\end{figure}

### 2017-08-16 Ceres/Old_versions/Altair_27_03_2015/CeresV4.tex
mod 2017 ago16

\title[Results of two stellar occultations by Ceres]{Results of two stellar occultations by dwarf planet (1) Ceres}
\author[A. V. Raveendran and A. N. Other]{A. V. Raveendran$^{1}$\thanks{E-mail:
email@address (AVR); otheremail@otheraddress (ANO)} and A. N.
Other$^{2}$\footnotemark[1]\thanks{This file has been amended to
highlight the proper use of \LaTeXe\ code with the class file.
These changes are for illustrative purposes and do not reflect the
original paper by A. V. Raveendran.}\\
$^{1}$Indian Institute of Astrophysics, Bangalore 560034, India\\
$^{2}$Building, Institute, Street Address, City, Code, Country}
\begin{document}

\date{Accepted 1988 December 15. Received 1988 December 14; in original form 1988 October 11}

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2002}

@@

\begin{keywords}
keywords.
\end{keywords}

@@

@@

Observations were carried out in Brazil at five stations as displayed in Table 1. The occultation was detected on four of them. Among these, one started the shots with the event already in progress and only provided star's reappearance time; the other three recorded the whole phenomenon.  Since this event had a very low velocity -- only 3.9 km/s -- the exposure times (which for faster occultations would result in a big time uncertainty) led to quite small error bars, with the exception of the UEPG chord. Moreover, as the larger timing errors occurred closer to the edges of the shadow's path, their radial-projected uncertainty came to be still smaller.

@@

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Circumstances of observation for all observing stations of the 2010 event.}
  \begin{tabular}{@{}lccccc}
  \hline
     Site & Longitude & Telescope & Exposure & Result & Observer  \\
          & Latitude  & Aperture  & Readout overhead  & Ingress & \\          
          & Height    & Camera    &    S/N    & Egress    & \\          
\hline
 Belo Horizonte & 43$\degr$59'51.1'' W & LX200 & 5 s & Positive & C. Jacques  \\
 CEAMIG-REA &19$\degr$49'49.0'' S & 31 cm &     & 22:39:03.9 $\pm$ 0.6 &  E. Pimentel \\
            & 825 m                &       &     & 22:40:20 $\pm$ 5 &   \\
 & & & & & \\
 Pico dos Dias    & 45$\degr$34'45.1'' W &       & 1 s & Positive & J. I. B. Camargo \\
 OPD/LNA    &22$\degr$32'03.7'' S & 60 cm &     & 22:37:30.3 $\pm$ 0.6 &  G. B. Rossi \\
            & 1864 m               &       &     & 22:41:55.3 $\pm$ 0.7 &              \\
 & & & & & \\
 S\~ao Jos\'e dos       & 45$\degr$51'44.0'' W &  C11  & 2 s & Egress only & A. C. Milone\\
 Campos       &23$\degr$12'33.0'' S & 28 cm &     & Start: 22:39:44 & T. Maldonado\\
 INPE      & 975 m                & SBIG ST7 &     & 22:42:03.0 $\pm$ 0.2 & M. Okada    \\
 & & & & & \\
 Ponta Grossa       & 50$\degr$05'56.0'' W & RCX 400 &30 s & Positive & M. Emilio   \\
 UEPG       &25$\degr$05'22.2'' S & 40 cm &     & 22:37:17 $\pm$ 13 & L. Mehret   \\
            & 910 m                & SBIG STL6E &     & 22:39:56 $\pm$ 13 & \\
 & & & & & \\
 Florian\'opolis       & 48$\degr$31'20.5'' W &       & 3 s & No occultation  & W. Schoenell\\
 UFSC       &27$\degr$36'12.3'' S & 28 cm &     &  & A. J. T. Mello\\
            & 20 m                 &       &     &         & F. R. Herpich \\
\hline
\end{tabular}
\end{minipage}
\end{table*}

[Curvas de luz, modelagem, difraç\~ao, di\^ametro estelar.]

@@

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Results of limb fitting to the data of the 2010 and 2013 events.}
  \begin{tabular}{@{}lcccc}
  \hline
     Solution & 2010/Nominal & 2010/Pole-constrained & 2013/Nominal & 2013/Pole-constrained \\
\hline
Equatorial radius (km) & 491 $\pm$ 7   & 486 $\pm$ 3   & 486 $\pm$ 2   & 485 $\pm$ 4\\
Oblateness             & 0.08 $\pm$ 0.03 & 0.08 $\pm$ 0.03 & 0.09 $\pm$ 0.05 & 0.08 $\pm$ 0.04\\
Position angle (deg)   & 5 $\pm$ 10    & 12 $\pm$ 3 (*)& 22 $\pm$ 5    &  25$\pm$3(*)\\
$f_c$ (km)             & 133 $\pm$ 9   & 138 $\pm$ 5   & 78 $\pm$ 7    & 78 $\pm$ 6\\
$g_c$ (km)             & -17 $\pm$ 15  & -11 $\pm$ 11  & 14 $\pm$ 16   & 13 $\pm$ 16\\
$\chi^2_{r,min}$       & 0.24          &  0.42         & 1.27          & 1.27\\
\hline
\end{tabular}
\textbf{Notes.} Error bars are at 1-sigma level. The reported oblateness derived from the 2010 event are the physical ones (not apparent), see text for details. \\
(*) Position angle derived from Ceres' rotational pole coordinates determined by \cite{Drummond2014}.
\end{minipage}
\end{table*}

\begin{figure}
\includegraphics[scale=0.36]{figures/Ceres_2010_body.pdf} 
\caption{\label{Fig:Ceres-2010-body}}
\end{figure}

@@

@@

\begin{figure}
\includegraphics[scale=0.42]{figures/Ceres_2013.png} 
\caption{Post-occultation reconstruction of Ceres' shadow path on Earth for the 2013 October 25 event at the east coast of USA. The red dots represents the center of the shadow separated by one minute where the shadow moves from the left to the right. The geocentric closest approach at 09:43:03 UT. The blue dots are the sites that have observed the event.\label{Fig: Ceres-2013-map}}
\end{figure}

@@

All videos were converted to FITS images and the flux of the star was obtained with the Platform for Reduction of Astronomical Images Automatically (PRAIA) \citep{2011gfun.conf...85A}. The light curves were normalized to an unocculted star flux when a star was available.

The data were binned by 5 or 10 points to reduce the noise. A linear or second-degree polynomial fit was applied to the observed flux before and after the occultation.

@@

The Fresnel scale ($F = \sqrt{\lambda D/2}$) for the present Ceres’ geocentric distance $D = 3.28 AU = 4.91 \times 10^{8} km$ is 0.4 km for a typical wavelength of $\lambda = 0.65 \micron m$.The star diameter is estimated using the formulae of \cite{vanBelle1999}. Its B, V , and K apparent magnitudes are 10.47, 9.97 and 8.76, respectively, in the Tycho catalog \citep{Hog2000}. This yields a diameter of about 0.16 km projected at Ceres’s distance. The smallest integration time used in the positive observations was 0.033 s, which translates to almost 1.4 km in the celestial plane. Therefore, the occultation light curves are largely dominated by the integration times, not by Fresnel diffraction or star diameter.

@@

\begin{figure}
\includegraphics[scale=0.58]{figures/Ceres_2013_fluxratio} 
\caption{Same as in Fig. \ref{Fig: Ceres-2010-curves} for the nini light curves of the 2013 event. \label{Fig: Ceres-2013-curves}}
\end{figure}

@@

From the occultation times obtained, the shape for Ceres' limb was adjusted by an ellipse characterized by five parameters: the coordinates of the body center, relative to the star in the plane of the sky ($f_{c}, g_{c}$); the apparent semimajor axis a, equivalent to the equatorial radius $R_{Equa}$ ; the apparent oblateness = $(a'-b')/a'$ (where $b'$ is the apparent semiminor axis or Polar Radius); and the position angle P of the semiminor axis $b'$ . The coordinates $f_{c}$ and $g_{c}$, in kilometers, are calculated using the JPL\#33  Ceres' ephemeris and the star position given in Section \ref{Sec: observation-2013}. They are positive toward the local celestial east and north, respectively.  The position angle P is counted positively from the direction of local celestial north to celestial east.

@@

\begin{figure}
\includegraphics[scale=0.36]{figures/Ceres_2013_body.pdf}
\caption{\label{Fig: Ceres-2013-body}}
\end{figure}

>>>>

\title[Results of two stellar occultations by Ceres]{Results of two stellar occultations by dwarf planet (1) Ceres}
\author[A. V. Raveendran and A. N. Other]{A. V. Raveendran$^{1}$\thanks{E-mail:
email@address (AVR); otheremail@otheraddress (ANO)} and A. N.
Other$^{2}$\footnotemark[1]\thanks{}\\
$^{1}$Observat\'orio do Valongo/UFRJ, Ladeira Pedro Ant\^onio 43,
CEP 20.080-090 Rio de Janeiro - RJ, Brazil\\
$^{2}$Observat\'orio Nacional/MCT, R. General Jos\'e Cristino 77,
CEP 20921-400 Rio de Janeiro - RJ, Brazil}
\begin{document}

\date{Accepted . Received ; in original form }

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2015}

@@

\begin{keywords}
minor planets, asteroids: individual (1, Ceres) - occultations - planets and satellites: fundamental parameters
\end{keywords}

@@

TYC 6833-163-1

@@

Observations were carried out in Brazil at five stations as displayed in Table 1. The occultation was detected on four of them. Among these, INPE started the shots with the event already in progress and only provided star's reappearance time; the other three recorded the whole phenomenon.  Since this event had a very low velocity -- only 3.9 km/s -- the exposure times (which for faster occultations would result in a big time uncertainty) led to quite small error bars, with the exception of the UEPG chord. Moreover, as the larger timing errors occurred closer to the edges of the shadow's path, their radial-projected uncertainty came to be still smaller.

@@

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Circumstances of observation for all observing stations of the 2010 event. \label{Tab: obs-2010}}
  \begin{tabular}{@{}lccccc}
  \hline
     Site & Longitude & Telescope & Exposure & Result & Observer  \\
          & Latitude  & Aperture  & Readout overhead  & Ingress & \\          
          & Height    & Camera    &    S/N    & Egress    & \\          
\hline
 Belo Horizonte & 43$\degr$59'51.1'' W & LX200 & 5 s & Positive & C. Jacques  \\
 CEAMIG-REA &19$\degr$49'49.0'' S & 31 cm &     & 22:39:03.9 $\pm$ 0.6 &  E. Pimentel \\
            & 825 m                &       &     & 22:40:20 $\pm$ 5 &   \\
 & & & & & \\
 Pico dos Dias    & 45$\degr$34'45.1'' W &       & 1 s & Positive & J. I. B. Camargo \\
 OPD/LNA    &22$\degr$32'03.7'' S & 60 cm &     & 22:37:30.3 $\pm$ 0.6 &  G. B. Rossi \\
            & 1864 m               &       &     & 22:41:55.3 $\pm$ 0.7 &              \\
 & & & & & \\
 S\~ao Jos\'e dos       & 45$\degr$51'44.0'' W &  C11  & 2 s & Egress only & A. C. Milone\\
 Campos       &23$\degr$12'33.0'' S & 28 cm &     & Start: 22:39:44 & T. Maldonado\\
 INPE      & 975 m                & SBIG ST7 &     & 22:42:03.0 $\pm$ 0.2 & M. Okada    \\
 & & & & & \\
 Ponta Grossa       & 50$\degr$05'56.0'' W & RCX 400 &30 s & Positive & M. Emilio   \\
 UEPG       &25$\degr$05'22.2'' S & 40 cm &     & 22:37:17 $\pm$ 13 & L. Mehret   \\
            & 910 m                & SBIG STL6E &     & 22:39:56 $\pm$ 13 & \\
 & & & & & \\
 Florian\'opolis       & 48$\degr$31'20.5'' W &       & 3 s & No occultation  & W. Schoenell\\
 UFSC       &27$\degr$36'12.3'' S & 28 cm &     &  & A. J. T. Mello\\
            & 20 m                 &       &     &         & F. R. Herpich \\
\hline
\end{tabular}
\end{minipage}
\end{table*}

\subsection{Light curves and timing \label{Sec: light-curve-2010}}

The flux of the star in the five occultation chords was obtained from the FITS images with the Platform for Reduction of Astronomical Images Automatically (PRAIA) \citep{2011gfun.conf...85A}. The light curves were normalized to an unocculted star flux. A linear or second-degree polynomial fit was applied to the observed flux before and after the occultation (with exception of INPE).

The Fresnel scale ($F = \sqrt{\lambda D/2}$) for the present Ceres’ geocentric distance $D = 2.29 AU = 3.42 \times 10^{8} km$ is 0.3 km for a typical wavelength of $\lambda = 0.65 \micron m$.The star diameter is estimated using the formulae of \cite{vanBelle1999}. Its B, V , and K apparent magnitudes are 12.00, 11.78 and 9.18, respectively, in the Tycho catalog \citep{Hog2000}. This yields a diameter of about $0.11 km$ projected at Ceres’s distance. The smallest integration time used in the positive observations was $1.0 s$, which translates to almost $3.9 km$ in the celestial plane. Therefore, the occultation light curves are largely dominated by the integration times, not by Fresnel diffraction or star diameter.

The best fits to the occultation light curves are shown in Figure \ref{Fig: Ceres-2010-curves}, and the derived occultation times are listed in Table \ref{Tab: obs-2010}.

@@

\begin{table*}
 \centering
 \begin{minipage}{140mm}
  \caption{Results of limb fitting to the data of the 2010 and 2013 events.}
  \begin{tabular}{@{}lcccc}
  \hline
     Solution & 2010/Nominal & 2010/Pole-constrained & 2013/Nominal & 2013/Pole-constrained \\
\hline
Equatorial radius (km) & 491 $\pm$ 7   & 486 $\pm$ 3   & 486 $\pm$ 4   & 485 $\pm$ 4\\
Oblateness             & 0.08 $\pm$ 0.03 & 0.08 $\pm$ 0.03 & 0.08 $\pm$ 0.04 & 0.08 $\pm$ 0.04\\
Position angle (deg)   & 5 $\pm$ 10    & 12 $\pm$ 3 (*)& 22 $\pm$ 5    &  25$\pm$3(*)\\
$f_c$ (km)             & 133 $\pm$ 9   & 138 $\pm$ 5   & 77 $\pm$ 6    & 78 $\pm$ 6\\
$g_c$ (km)             & -17 $\pm$ 15  & -11 $\pm$ 11  & 13 $\pm$ 16   & 13 $\pm$ 16\\
$\chi^2_{r,min}$       & 0.24          &  0.42         & 1.27          & 1.27\\
\hline
\end{tabular}
\textbf{Notes.} Error bars are at 1-sigma level. The reported oblateness derived from the 2010 event are the physical ones (not apparent), see text for details. \\
(*) Position angle derived from Ceres' rotational pole coordinates determined by \cite{Drummond2014}.
\end{minipage}
\end{table*}

\begin{figure}
\includegraphics[scale=0.36]{figures/Ceres_2010_body.pdf} 
\caption{The best elliptical fit for the occultation chords for the event of 2010 using timing from Table \ref{Tab: obs-2010} and the pole-constrained solution. \label{Fig:Ceres-2010-body}}
\end{figure}

@@

TYC 865-911-1

@@

\begin{figure}
\includegraphics[scale=0.42]{figures/Ceres_2013.png} 
\caption{Post-occultation reconstruction of Ceres' shadow path on Earth for the 2013 October 25 event at the east coast of USA. The big dot is the geocentric closest approach at 09:43:03 UT and the small ones represents the center of the shadow separated by 30 seconds where the shadow moves from the left to the right.  The blue dots are the sites that have observed the event.\label{Fig: Ceres-2013-map}}
\end{figure}

@@

All videos were converted to FITS images and the flux of the star was also obtained with the PRAIA \citep{2011gfun.conf...85A}. The light curves were normalized to an unocculted star flux when a star was available.

@@

Due to the small integration time, the flux star was very noisy. To reduce the noise, the data were binned by 5 points for all curves, except for Greenbelt where a 10 points binning was needed. This made an effective integration time 5 or 10 times bigger.

The star and end times were obtained by the same procedure explained in Section \ref{Sec: light-curve-2010}. The Fresnel scale of $0.4 km$ for this event was calculated using the Ceres’ geocentric distance $D = 3.28 AU = 4.91 \times 10^{8} km$and a typical wavelength of $\lambda = 0.65 \micron m$.The star diameter projected at Ceres' distance of $0.16 km$ is estimated using the B = 10.47, V=9.97, and K=8.76 apparent magnitudes in the Tycho catalog \citep{Hog2000}. The smallest effective integration time used in the positive observations was $0.165 s$, which translates to almost $7.0 km$ in the celestial plane. As for the 2010 event, the occultation light curves are largely dominated by the integration times, not by Fresnel diffraction or star diameter.

@@

\begin{figure}
\includegraphics[scale=0.58]{figures/Ceres_2013_fluxratio} 
\caption{The nine occultation light curves normalized and vertically shifted by a factor of 1.0 for better viewing. The solid black lines are the best fit of the square-well model to the data. The red lines are the square-well model convoluted with the Fresnel diffraction, the star diameter, and the applied exposure time. The mid-times of the occultations do not coincide due to the propagation delays of the shadow due to the distinct longitude of the sites. The light curve of Brookline is shifted by -64s as explained in the text. \label{Fig: Ceres-2013-curves}}
\end{figure}

@@

From the occultation times obtained, the shape for Ceres' limb was adjusted by an ellipse characterized by five parameters: the coordinates of the body center, relative to the star in the plane of the sky ($f_{c}, g_{c}$); the apparent semimajor axis $a'$, equivalent to the equatorial radius $R_{Equa}$ ; the apparent oblateness = $(a'-b')/a'$ (where $b'$ is the apparent semiminor axis); and the position angle P of the semiminor axis $b'$ . The coordinates $f_{c}$ and $g_{c}$, in kilometers, are calculated using the JPL\#33  Ceres' ephemeris and the star position given in Section \ref{Sec: observation-2013}. They are positive toward the local celestial east and north, respectively.  The position angle P is counted positively from the direction of local celestial north to celestial east.

@@

\begin{figure}
\includegraphics[scale=0.36]{figures/Ceres_2013_body.pdf}
\caption{The best elliptical fit for the occultation chords for the event of 2013 using timing from Table \ref{Tab: obs-2013} and the pole-constrained solution. The chord of Brookline if shifted by -64s as explained in the text\label{Fig: Ceres-2013-body}}
\end{figure}
